---
title: "NLoN - Natural Language or Not"
author: "Maëlick Claes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NLoN - Natural Language or Not}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette presents how one can train a model with NLoN and use to
predict whether lines of text are natural language or not. It also
shows how to use a different training dataset than the one provided in
this package, and how to evaluate its performance.

## Data preparation

This paper includes the training dataset that was used in our MSR
paper "Natural Language or Not (NLoN) - A Package for Software
Engineering Text Analysis
Pipeline" <https://arxiv.org/abs/1803.07292>. The dataset is presented
as a *data.table* object with different columns:

* *source* is a character string containing the source of the line of
  text (*mozilla*, *kubernetes*, and *lucene*).
* text is a character string containing the line of text
* rater1 and rater2 are the factor (with levels *NL* for natural
  language and *Not* for anything else than natural language) with the
  manual labeling of the data done by two different persons.

```{r}
library(data.table)
library(NLoN)

nlon.data
```

In order to train a model, only a vector with the text and a
vector (of the same length) with the factor response are needed. A
source vector was added in order to identify where does each line of
text provided in the package come from.

## Computing features

Before training a model, it is necessary to build features that will
be used by *glmnet*. Various functions are defined in the module
*NLoN::features* to compute individual features. For example

```{r}
text <- c("This is some text.", "This contains 1 number.", "123")
NLoN::features$Numbers(text)
```

The package also includes three other functions computing multiple
features at the same time. *NLoN::FeatureExtraction* computes the set
of simple text features that were used in the NLoN MSR paper and were
inspired by regular expression matching. *NLoN::Character3Grams*
returns a document term sparse matrix with the character tri-grams
contained the the various lines of text. Finally
*NLoN::TriGramsAndFeatures* returns a matrix with both the two other
functions.

## Training a model

A model can be trained using *NLoN::NLoNModel*. A model can simply be
trained by providing the text and expected response (a factor with
levels *NL* and *Not*).

```{r, message='hide', fig.width=7, fig.height=4}
model <- NLoN::NLoNModel(nlon.data$text, nlon.data$rater2)
plot(model)
```

By default features will be computed using *NLoN::TriGramsAndFeatures*
and a 10-fold cross validation. Several optional arguments allow to
customize the model.

A different feature function can be specified using the *features*
argument. The function must return either a matrix, a data.frame with
numerical values or a list of numerical vectors of the same
length. Alternatively, features can also be a list of functions
returning individual numeric vectors.

The cross validation can also be customized. By default it uses *glmnet*
as it is faster, but *caret* can also be used to run a repeated
10-fold cross validation. This is done by changing the *cv* to
*"caret"*. The number of repeats is 10 by default but can be changed
with the parameter *repeats*. (Note that cross validation with *caret*
is significantly slower than with *glmnet*.) Finally if a value different than
*"caret"* or *"glmnet"* is passed to *cv*, then no cross validation is
used and a single model is trained on the whole dataset.

For example we can build a model without using character tri-grams:

```{r, fig.width=7, fig.height=4}
model2 <- NLoN::NLoNModel(nlon.data$text, nlon.data$rater2,
                          features=FeatureExtraction)
plot(model2)
```

Or build one using only the ratio of uppercase letters, numbers
and special characters and a cross-validation repeated 5 times:

```{r, fig.width=7, fig.height=4}
features <- list(caps.ratio=NLoN::features$CapsRatio,
                 numbers.ratio=NLoN::features$NumbersRatio,
                 special.ratio=NLoN::features$SpecialCharsRatio)
system.time(model3 <- NLoN::NLoNModel(nlon.data$text, nlon.data$rater2,
                                      features=features, cv="caret", repeats=5,
                                      verbose=FALSE))
plot(model3)
```

## Making predictions

Once a model has been built, *NLoN::NLoNPredict* can be used to make
predictions. The function requires an object as returned by
*NLoN::NLoNModel* and a character vector containing the text on which
make the predictions:

```{r}
text <- c("This is natural language.", "not(natural, language);")
NLoN::NLoNPredict(model, text)
```

By default the prediction is done using *NLoN::TriGramsAndFeatures*,
but if the model was trained differently, the feature function must
also be passed to *NLoN::NLoNPredict*:

```{r}
NLoN::NLoNPredict(model, text, features=FeatureExtraction)
```

By default the prediction is done at minimum lambda (*lambda.min*),
but it can also be done at *lambda.1se* which gives the most
regularized (penalized) model such that AUC is within one standard
deviation of the maximum. Note that *lambda.min* is only available for
cross validated models (either trained with *caret* or *glmnet*) and
*lambda.1se* for cross validated models trained with *glmnet*.

If no cross validation was done before, lambda needs to be specified
manually. Not performing cross validation can be a way to save time
when training the model. Repeated cross validation (with *caret*) can be
particularly slow. Thus it can be a good idea to run the cross
validation only the first time, save the value of *lambda.min* and
train the model without cross validation for future use.

For example we trained with *caret* a full model with a 10 repeated 10
fold cross validation, which takes 1850 seconds to complete. We
obtained a *lambda.min* value of 0.003448735. We can now train a model
on the full training set in a very short time and directly used the
lambda value previously obtained with cross validation:

```{r}
lambda.min <- 0.003784982
system.time(model4 <- NLoN::NLoNModel(nlon.data$text, nlon.data$rater2,
                                      cv="none"))
NLoN::NLoNPredict(model4, text, lambda=lambda.min)
```

## Evaluating models

### All sources

The first model trained on all sources with cross validation using
both 3-grams and feature engineering has the following AUC at
*lambda.min*:

```{r}
with(model, cvm[lambda == lambda.min])
```

And the following AUC at *lambda.1se*:

```{r}
with(model, cvm[lambda == lambda.1se])
```

### Within-source prediction

Here we train one model for each source.

### Mozilla

```{r, fig.width=7, fig.height=4}
comments.model <- NLoN::NLoNModel(nlon.data[source == "mozilla", text],
                                  nlon.data[source == "mozilla", rater2])
plot(comments.model)
with(comments.model, cvm[lambda == lambda.min])
with(comments.model, cvm[lambda == lambda.1se])
```

### Lucene

```{r, fig.width=7, fig.height=4}
emails.model <- NLoN::NLoNModel(nlon.data[source == "mozilla", text],
                                  nlon.data[source == "mozilla", rater2])
plot(emails.model)
with(emails.model, cvm[lambda == lambda.min])
with(emails.model, cvm[lambda == lambda.1se])
```

### Kubernetes

```{r, fig.width=7, fig.height=4}
chats.model <- NLoN::NLoNModel(nlon.data[source == "mozilla", text],
                                  nlon.data[source == "mozilla", rater2])
plot(chats.model)
with(chats.model, cvm[lambda == lambda.min])
with(chats.model, cvm[lambda == lambda.1se])
```

### Cross-source prediction

Here we test how well a single source can be predicted using the two
other data sources to train a model.

```{r}
mozilla.train <- nlon.data[source != "mozilla"]
mozilla.test <- nlon.data[source == "mozilla"]
mozilla.model <- NLoN::NLoNModel(mozilla.train$text, mozilla.train$rater2)
mozilla.resp <- NLoN::NLoNPredict(mozilla.model, mozilla.test$text,
                                  type="response")
mozilla.predict <- NLoN::NLoNPredict(mozilla.model, mozilla.test$text)
ModelMetrics::auc(mozilla.test$rater2, mozilla.resp)
table(mozilla.predict, mozilla.test$rater2)
```

```{r}
kubernetes.train <- nlon.data[source != "kubernetes"]
kubernetes.test <- nlon.data[source == "kubernetes"]
kubernetes.model <- NLoN::NLoNModel(kubernetes.train$text, kubernetes.train$rater2)
kubernetes.resp <- NLoN::NLoNPredict(kubernetes.model, kubernetes.test$text,
                                  type="response")
kubernetes.predict <- NLoN::NLoNPredict(kubernetes.model, kubernetes.test$text)
ModelMetrics::auc(kubernetes.test$rater2, kubernetes.resp)
table(kubernetes.predict, kubernetes.test$rater2)
```

```{r}
lucene.train <- nlon.data[source != "lucene"]
lucene.test <- nlon.data[source == "lucene"]
lucene.model <- NLoN::NLoNModel(lucene.train$text, lucene.train$rater2)
lucene.resp <- NLoN::NLoNPredict(lucene.model, lucene.test$text,
                                  type="response")
lucene.predict <- NLoN::NLoNPredict(lucene.model, lucene.test$text)
ModelMetrics::auc(lucene.test$rater2, lucene.resp)
table(lucene.predict, lucene.test$rater2)
```
